---
title: "Milestone_3"
author: "Martha Wetzel"
date: "4/26/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


#library(foreign)  
library(tidyverse)
#library(haven)
library("sas7bdat")
library(ggplot2)
library(kableExtra)
library(dplyr)
#library(survey)
#library(srvyr)
library(car)
library(stargazer)
library(broom)


load(file = "mydata3.RData")

```


```{r missings}

alldata$Complete <- (is.na(alldata$Employed) == F & is.na(alldata$DaysLimited) == F)
alldata$Complete <- complete.cases(alldata)


```


# Results
## Descriptive Statistics

The tables below provide descriptive statistics for key demographic variables.

```{r descriptives} 

# Create factor variable for policy type
alldata <- alldata %>% mutate(CurrentPolicy = if_else(PostMandPDMP ==1, 3, 
                                                                       if_else( PostOptPDMP == 1, 2, 1)))

knitr::kable(x = table(alldata$FEMALE, alldata$CurrentPolicy),
             col.names = c("No PDMP", "Optional PDMP", "Mandatory PDMP"),
             caption = "Gender By PDMP Policy")

# General Health

knitr::kable(x = table(alldata$GenHealthFac, alldata$CurrentPolicy),
             col.names = c("No PDMP", "Optional PDMP", "Mandatory PDMP"),
             caption = "Health Status By PDMP Policy",
             row.names = TRUE)

# Education

knitr::kable(x = table(alldata$EDUCA_cln, alldata$CurrentPolicy),
             col.names = c("No PDMP", "Optional PDMP", "Mandatory PDMP"),
             caption = "Education By PDMP Policy",
             row.names = TRUE)

# Employment

knitr::kable(x = table(alldata$Employed, alldata$CurrentPolicy),
             col.names = c("No PDMP", "Optional PDMP", "Mandatory PDMP"),
             caption = "Employment By PDMP Policy",
             row.names = TRUE)


```

The following figures illustrate the distribution of focal outcomes. Both the outcomes of days where physical health was not good and days where activity was limited due to poor health had a range of 0 - 30. Figures 1 and 2 show days of not good physical health. The distribution was similar between the optional and mandatory PDMPs. In both cases, the distribution was bimodal. Due to this distribution, outcomes were recoded as binary variables indicating the individual spent more than half of the preceeding 30 days in poor physical health.

```{r histogramsphy}
# Illustrate the outcomes

alldata <- alldata %>% mutate(OptPDMPLabel = if_else(PostOptPDMP==0, "No Optional PDMP", "Optional PDMP"))
alldata <- alldata %>% mutate(MandPDMPLabel = if_else(PostMandPDMP==0, "No Mandatory PDMP", "Mandatory PDMP"))

# Physical Health not good
# Optional PDMP Distribution
ggplot(alldata, aes(PhysHealthBad, fill = OptPDMPLabel)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'dodge') + 
  ggtitle("Poor Physical Health \n By Optional PDMP Status") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Days", y = "Density", fill="PDMP Status")

# Mandatory PMDP Distribution
ggplot(alldata, aes(PhysHealthBad, fill = MandPDMPLabel)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'dodge') + 
  ggtitle("Poor Physical Health \n By Mandatory PDMP Status") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Days", y = "Density", fill="PDMP Status")

```

Similar to the days of poor physical health, the days of activity limitations outcome also showed a bimodal distribution. There was a small difference visible for the mandatory vs. not mandatory PDMP distribution, but overall the distributions were very similar. This variable was also recoded to be a binary indicator representing individuals who reported having activity limitations for more than half the days of the month.


```{r histogramslim}
# Days Limited
# Optional PDMP Distribution
ggplot(alldata, aes(DaysLimited, fill = OptPDMPLabel)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'dodge') + 
  ggtitle("Days of Activity Limitations \n By Optional PDMP Status") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Days", y = "Density", fill="PDMP Status")

# Optional PDMP Distribution
ggplot(alldata, aes(DaysLimited, fill = MandPDMPLabel)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'dodge') + 
  ggtitle("Days of Activity Limitations \n By Mandatory PDMP Status") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Days", y = "Density", fill="PDMP Status")
```

## Regression Results


```{r}


# Start with employment, most straight forward for replication

Emp_reg <- glm(alldata$Employed ~ PostOptPDMP + PostMandPDMP + GenHealthFac + MEDCOST_cln + Race_cln +
                      FEMALE + Age  +  EDUCA_cln + SMOKE100_cln + year_char + Statecode,
               data=alldata, family=binomial)

# Get odds ratios and transform SE 
Emp_reg_df <- tidy(Emp_reg)

# this is being done the long way because mutate seems to be broken 
Emp_reg_df$or <- exp(Emp_reg_df$estimate)
Emp_reg_df$var_diag <- diag(vcov(Emp_reg))
Emp_reg_df$or_se <- sqrt(Emp_reg_df$or^2 * Emp_reg_df$var_diag)
Emp_reg_df$stars <-  if_else(Emp_reg_df$p.value < 0.01, "***", 
                          if_else (Emp_reg_df$p.value < 0.05, "**",
                          if_else(Emp_reg_df$p.value < 0.1, "*", " ") )) # Add stars by hand

# Clean it up for printing
Emp_reg_df2 <- select(Emp_reg_df, term, or, or_se, stars)
Emp_reg_df2$or <- round(Emp_reg_df2$or, 3)
Emp_reg_df2$or_se <- round(Emp_reg_df2$or_se, 3)
Emp_reg_df2

```

```{r physhealthreg}
# Physical Health Bad More Than Half the Month

PhysBad <- glm(alldata$BadPhys_Bin ~ PostOptPDMP + PostMandPDMP + GenHealthFac + MEDCOST_cln + Race_cln + FEMALE
               + Age  + 
                 EDUCA_cln + SMOKE100_cln + year_char + Statecode,
               data=alldata, family=binomial)


# Get odds ratios and transform SE 
PhysBad_df <- tidy(PhysBad)

# this is being done the long way because mutate seems to be broken 
PhysBad_df$or <- exp(PhysBad_df$estimate)
PhysBad_df$var_diag <- diag(vcov(PhysBad))
PhysBad_df$or_se <- sqrt(PhysBad_df$or^2 * PhysBad_df$var_diag)
PhysBad_df$stars <-  if_else(PhysBad_df$p.value < 0.01, "***", 
                          if_else (PhysBad_df$p.value < 0.05, "**",
                          if_else(PhysBad_df$p.value < 0.1, "*", " ") )) # Add stars by hand

# Clean it up for printing
PhysBad_df2 <- select(PhysBad_df, term, or, or_se, stars)
PhysBad_df2$or <- round(PhysBad_df2$or, 3)
PhysBad_df2$or_se <- round(PhysBad_df2$or_se, 3)
PhysBad_df2
```

```{r limitreg}
# Days Limited
Limited <- glm(alldata$Limited_bin ~ PostOptPDMP + PostMandPDMP + GenHealthFac + MEDCOST_cln + Race_cln + FEMALE
               + Age  +  EDUCA_cln + SMOKE100_cln + year_char + Statecode,
               data=alldata, family=binomial)


models <- list(PhysBad, Limited)


# Get odds ratios and transform SE 
PhysBad_df <- tidy(PhysBad)

# this is being done the long way because mutate seems to be broken 
PhysBad_df$or <- exp(PhysBad_df$estimate)
PhysBad_df$var_diag <- diag(vcov(PhysBad))
PhysBad_df$or_se <- sqrt(PhysBad_df$or^2 * PhysBad_df$var_diag)
PhysBad_df$stars <-  if_else(PhysBad_df$p.value < 0.01, "***", 
                          if_else (PhysBad_df$p.value < 0.05, "**",
                          if_else(PhysBad_df$p.value < 0.1, "*", " ") )) # Add stars by hand

# Clean it up for printing
PhysBad_df2 <- select(PhysBad_df, term, or, or_se, stars)
PhysBad_df2$or <- round(PhysBad_df2$or, 3)
PhysBad_df2$or_se <- round(PhysBad_df2$or_se, 3)
PhysBad_df2
```




